---
title: "EPiC_23a Analysis Script"
author: "Nicholas Coles"
date: "`r Sys.Date()`"
output: word_document
editor_options: 
  chunk_output_type: console
---

# Set environment
```{r}
# clear environment
rm(list = ls())

# install (if necessary) and load packages
# function written by stevenworthington 
Ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, 'Package'])]
  
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  
  sapply(pkg, require, character.only = TRUE)
}

# vector of necessary packages
packages <- c('tidyverse', 'lme4',
              'lmerTest', 'emmeans',
              'gtools')

# using vector of packages, call ipak function
Ipak(packages)

# delete vestigial
rm(packages, Ipak)

# turn scientific notation off
options(scipen = 999)

# set seed to year of Nicholas' favorite [unfinished] album, SMiLE
set.seed(1967)

# allow for Type III Sums of Square
options(contrasts = c('contr.sum', 'contr.poly'))

# set plotting theme
theme_set(theme_classic())

# free up computational resources
gc()
memory.limit(size = 25000)
```

# Load data
```{r}
DF <- readRDS('DF.rds') 
```

# Create the mega-model
```{r eval = F}
tmp <- DF %>% 
  filter(team != "team 4 (random)",
         team != "team 9 (random)",
         team != "team 3 (random)",
         team != "baseline_tailored") %>% 
  
  # extract info
  mutate(source = 
           if_else(condition = grepl("team", team),
                   true = "team",
                   false = "baseline"),
         team.name = 
           if_else(source == "team",
                   true = team,
                   false = NA),
         baseline.type = 
           if_else(team == "baseline_random",
                   true = "random",
                   false = 
                     if_else(team == "baseline_mean",
                             true = "mean",
                             false = NA)
           )
  ) %>% 
  
  # refactor
  mutate(source = factor(source),
         team.name = factor(team.name),
         baseline.type = factor(baseline.type))

mega.m <- 
  lmer(error ~ source : baseline.type : team.name : scenario +
         (1 | sub) + (1 | vid),
       data = tmp)
```


# Examine error in  teams' submission
Create MainAnalyses function
```{r}
MainAnalyses <- function(t){
  # filter data
  d <- DF %>% 
    filter(team == "baseline_random" |
             team == "baseline_mean" |
             team == t) %>%
    #group_by(scenario) %>% 
    #sample_n(size = 1000) %>% 
    #ungroup() %>% 
    mutate(team = factor(team))
  
  # fit model
  m <- lmer(error ~ scenario * team +
              (1 | sub_orig) +
              (1 | vid_orig),
            data = d)
  
  # run ANOVA and perform contrasts
  # m.aov <- m %>% 
  #   anova() %>% 
  #   mutate(team = t)
  
  m.emm <- m %>% 
    emmeans(pairwise ~ team | scenario) %>% 
    .$contrasts %>% 
    as.data.frame %>% 
    filter(grepl(t, contrast)) %>% 
    mutate(team = t, 
           contrast = str_sub(contrast,
                              start = 9, end = 13))
  
  # compile results into list
  return(m.emm)
}
```

Run function
```{r}
t.list <- c("team 1", "team 2", "team 3",
            "team 4", "team 5", "team 6",
            "team 7", "team 8", "team 9",
            "team 10", "team 11", "team 12")

results <- sapply(FUN = MainAnalyses,
                  X = t.list,
                  simplify = F)

rm(t, t.list, d, m, m.emm)

saveRDS(results, 'results.Rds')
```

# Figure 0a
This figure could be cool because it could demonstrate that the test data has the same differentiation as the Sharma dataset.

Could be more trouble than it's worth though.

You could imagine this being the A part of a spirit figure.

Particpiants encountered 4 inductions of two affective states while providing moment-to-moment measures of valence and arousal

Data are split for four types of validation

```{r}
# add video label
DF <- DF %>% 
  mutate(`targeted emotion` = 
           if_else(
             vid_orig == 1 | 
               vid_orig == 2,
             true = "amusing",
             false = 
               if_else(
                 vid_orig == 3 |
                   vid_orig == 4,
                 true = "boring",
                 false = 
                   if_else(
                     vid_orig == 5 |
                       vid_orig == 6,
                     true = "relaxing",
                     false = "scary")
               )
             )
         )

# create participant level summaries
f0.dfa <- DF %>% 
  filter(team == "baseline_mean",
         scenario == "scenario_4") %>% 
  select(sub_orig, vid_orig, `targeted emotion`, time, scenario, fold, outcome, test) %>% 
  pivot_wider(names_from = 'outcome',
              values_from = 'test') %>% 
  group_by(sub_orig, `targeted emotion`) %>% 
  summarise(m.val = mean(valence),
            m.aro = mean(arousal))

# create video level summaries
f0.dfb <- DF %>% 
  filter(team == "baseline_mean",
         scenario == "scenario_4") %>% 
  select(sub_orig, vid_orig, `targeted emotion`, time, scenario, fold, outcome, test) %>% 
  pivot_wider(names_from = 'outcome',
              values_from = 'test') %>% 
  group_by(`targeted emotion`) %>% 
  summarise(m.val = mean(valence),
            sd.val = sd(valence),
            m.aro = mean(arousal),
            sd.aro = sd(arousal))

ggplot(data = f0.dfa,
       aes(x = m.val,
           y = m.aro,
           fill = `targeted emotion`
           )
       ) +
  stat_ellipse(geom = "polygon",
               alpha = .2) +
  geom_point(aes(color = `targeted emotion`)) +
  ylab('mean valence reports') +
  xlab('mean arousal reports')
```

# Figure 0b


# Old Figure 1
```{r}
#results <- readRDS('results.Rds')

# create dataframe containing contrasts
DF.f1.contrasts <- bind_rows(results) %>%
  select(contrast:estimate, z.ratio, p.value, team) %>%
  mutate(p.value = stars.pval(p.value),
         estimate = round(estimate, 2)) %>%
  pivot_wider(names_from = contrast,
              values_from = c(estimate, z.ratio, p.value))

# create dataframe containing baseline performance
DF.f1.baselines <- DF %>%
  filter(team == "baseline_mean" |
           team == "baseline_random") %>%
  group_by(team, scenario) %>%
  summarise(m.err = mean(error)) %>%
  pivot_wider(names_from = team,
              values_from = m.err)

# create dataframe containing team performance
DF.f1.teams <- DF %>%
  filter(
    team != "team 3 (random)",
    team != "team 4 (random)",
    team != "team 9 (random)",
    team != "baseline_mean",
    team != "baseline_random",
    team != "baseline_tailored"
  ) %>%
  group_by(team, scenario) %>%
  summarise(m.err = mean(error))

# merge dataframes
DF.f1 <-
  full_join(x = DF.f1.contrasts,
            y = DF.f1.baselines) %>%
  full_join(x = .,
            y = DF.f1.teams) %>%
  mutate(
    scenario =
      factor(
        scenario,
        labels =
          c(
            'across-time validation',
            'across-subject validation',
            'across-emotion validation' ,
            'across-induction validation'
          )
      ),
    scenario =
      factor(
        scenario,
        levels = c(
          'across-subject validation',
          'across-emotion validation' ,
          'across-induction validation',
          'across-time validation'
        )
      ),
    team = factor(team,
                  levels = paste0('team ',
                                  seq(12, 1)))
  )

rm(DF.f1.baselines, DF.f1.contrasts, DF.f1.teams)
```

Summary statistics
```{r}
# results from baseline__rand
DF.f1 %>%
  filter(!is.na(estimate__rand)) %>%
  summarise(
    min.diff = min(estimate__rand),
    max.diff = max(estimate__rand),
    min.z = min(z.ratio),
    max.z = max(z.ratio)
  )

# results from baseline__mean
DF.f1 %>%
  filter(estimate__mean > 0) %>%
  group_by(scenario) %>%
  summarise(
    min.diff = min(estimate__mean),
    max.diff = max(estimate__mean),
    min.z = min(z.ratio),
    max.z = max(z.ratio)
  ) %>%
  View()

DF.f1 %>%
  filter(estimate__mean > 0,
         scenario == "across-time validation") %>%
  View()

# how often did models beat baseline
tmp <- DF.f1 %>%
  mutate(
    beat_mean =
      if_else(
        estimate__mean > 0 &
          p.value__mean != " ",
        true = "yes",
        false = "no"
      ),
    
    beat_rand =
      if_else(
        estimate__rand > 0 &
          p.value__rand != " ",
        true = "yes",
        false = "no"
      )
  )

tmp$beat_rand %>% table %>% prop.table

tmp$beat_mean %>% table %>% prop.table

# what is overall accuracy of each model
tmp <- DF %>%
  group_by(team) %>%
  summarise(m.err = mean(error)) %>%
  arrange(m.err)

# team 2 and 10 are similar (worst similar pair)
# for team 2, across-subject and across-time were similar
# for team 10, across-time was much worse
tmp.2.10 <- DF.f1 %>%
  filter(team == "team 2" | team == "team 10") %>%
  arrange(team, m.err)

# team 6 and 12 (better similar pair)
# similar inferences about across-subject vs. across-time. Opposite inferences about across-induction and across-time
tmp.6.12 <- DF.f1 %>%
  filter(team == "team 6" | team == "team 12") %>%
  arrange(team, m.err)

# team 9 and 11 (best pair)
# opposite inferences about across-time vs. across-subject
tmp.9.11 <- DF.f1 %>%
  filter(team == "team 9" | team == "team 11") %>%
  arrange(team, m.err)
```

Plot figure
```{r}
ggplot(DF.f1,
       aes(x = m.err,
           y = team)) +
  facet_wrap( ~ scenario,
              nrow = 4,
              ncol = 1) +
  
  # add team performance bars
  geom_bar(stat = "identity",
           fill = '#C0C0F0') +
  
  # add lines for baseline_random and baseline_mean performance
  geom_vline(linetype = 'dotted',
             aes(xintercept = baseline_random),
             colour = '#3366FF') +
  geom_vline(linetype = 'dashed',
             aes(xintercept = baseline_mean),
             colour = '#3366FF') +
  
  # add summary statistic text
  geom_text(aes(
    x = baseline_random + .15,
    label = paste0(estimate__rand %>%
                     format(nsmall = 2),
                   p.value__rand)
  ),
  size = 3,
  hjust = 0) +
  geom_text(aes(
    x = baseline_mean + .15,
    label = paste0(estimate__mean %>%
                     format(nsmall = 2),
                   p.value__mean)
  ),
  size = 3,
  hjust = 0) +
  ylab('model developers') +
  xlab('mean absolute prediction error') +
  scale_x_continuous(expand =
                       expansion(mult = c(0, .05)))
```



# New idea for Figure 1

```{r}
tmp <- DF %>% 
  group_by(team, scenario) %>% 
  summarise(m.err = mean(error)) %>% 
  filter(team != "team 9 (random)",
         team != "team 4 (random)",
         team != "team 3 (random)") %>% 
  mutate(scenario = 
           factor(scenario,
                  labels = 
                    c('across-time',
                      'across-subject',
                      'across-emotion' ,
                      'across-induction')),
         scenario =
           factor(scenario,
                  levels = c('across-subject',
                             'across-emotion' ,
                             'across-induction',
                             'across-time'))
  )

# ggplot(tmp,
#        aes(x = scenario,
#            y = m.err,
#            group = team,
#            colour = team)) +
#   stat_summary(geom = "line",
#                size = 1) +
#   scale_y_reverse() +
#   scale_x_discrete(position = "top") 
```


##  flipped spaghetti                        
```{r}
g <- ggplot(tmp,
            aes(y = scenario,
                x = m.err,
                group = team,
                colour = team)) +
  
  # create baseline areas
  geom_area(data = filter(tmp, 
                          str_detect(team, "baseline_random")), 
            colour = NA,
            fill = "gray85") +
  geom_area(data = filter(tmp, 
                          str_detect(team, "baseline_mean")), 
            colour = NA,
            fill = "grey95") +
  
  # label  baseline areas
  ## baseline_mean
  annotate(geom = "line",
           y = c(0 + .02, 
                 tmp[tmp$team == "baseline_mean" &
                       tmp$scenario == "across-subject", ]$m.err - .02),
           x = .75,
           size = 1,
           colour = "grey60"
           ) +
  
  annotate(geom = "text",
           y = tmp[tmp$team == "baseline_mean" &
                     tmp$scenario == "across-subject", ]$m.err / 2,
           x = .75,
           label = "above relying on sample averages",
           colour = "grey60",
           angle = 90,
           vjust = -1)

g <- g +
  ## baseline_random
  annotate(geom = "line",
           y = c(0,
                 # tmp[tmp$team == "baseline_mean" &
                 #       tmp$scenario == "across-subject", ]$m.err + .02, 
                 tmp[tmp$team == "baseline_random" &
                       tmp$scenario == "across-subject", ]$m.err - .02),
           x = .9,
           size = 1,
           colour = "grey40") +
  
  annotate(geom = "text",
           y = 2,
           x = .75,
           label = "above chance",
           angle = 90,
           vjust = -1,
           colour = "grey40") +
  
  # add teams
  geom_line(data = filter(tmp, 
                          !str_detect(team, "baseline")), 
               alpha = .5, 
            size = 2) +
  
  ggrepel::geom_text_repel(data = filter(tmp, 
                                         !str_detect(team, "baseline"), 
                                         scenario == "across-time"), 
                           aes(label = team, 
                               col = team),
                           hjust = 1,
                           nudge_x = 1,
                           force = 2) + 
  scale_x_discrete(expand = expansion(mult = c(.2, .2))) +
  #scale_y_continuous(expand = expansion(mult = c(0, 0))) +
  ylab('mean absolute prediction error') +
  scale_y_reverse(expand = expansion(mult = c(0, 0))) +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  
  ggrepel::geom_text_repel(data = filter(tmp, 
                                         team == "team 5",
                                         scenario == "across-subject"),
                           aes(label = "italic('more')~accurate~across~subjects~than~time",
                               col = team),
                           nudge_y = .3,
                           nudge_x = .9,
                           parse = T) +
  
  ggrepel::geom_text_repel(data = filter(tmp, 
                                         team == "team 8",
                                         scenario == "across-subject"),
                           aes(label = "italic('less')~accurate~across~subjects~than~time",
                               col = team),
                           nudge_y = -.3,
                           nudge_x = .9,
                           parse = T)
```


Next steps for this:
* Annotate Team 5 and 8
* Higher accuracy than mean baseline model (doesn't look great)
* Higher accuracy than random baseline model (doesn't look great)
* Re-consider ordering of teams

Thought about figure: we could hypothetically show people the distribution of the data set


Thought:
Red team.

The models appear to be capturing something. Whether it's a true link with physiology, we cannot say for sure.


Core affect may not merely be physiological

Are links between physiology and core affect 

##  Mikes's code                       
```{r}
g <- ggplot(tmp,
       aes(x = scenario,
           y = m.err,
           group = team,
           colour = team)) +
  
  # create baseline areas
  geom_area(data = filter(tmp, 
                          str_detect(team, "baseline_random")), 
            colour = NA,
            fill = "gray85") +
  geom_area(data = filter(tmp, 
                          str_detect(team, "baseline_mean")), 
            colour = NA,
            fill = "grey95") +
  
  # label  baseline areas
  ## baseline_mean
  annotate(geom = "line",
           y = c(0 + .02, 
                 tmp[tmp$team == "baseline_mean" &
                       tmp$scenario == "across-subject", ]$m.err - .02),
           x = .75,
           size = 1,
           colour = "grey60"
           ) +
  
  annotate(geom = "text",
           y = tmp[tmp$team == "baseline_mean" &
                     tmp$scenario == "across-subject", ]$m.err / 2,
           x = .75,
           label = "above relying on sample averages",
           colour = "grey60",
           angle = 90,
           vjust = -1)

g <- g +
  ## baseline_random
  annotate(geom = "line",
           y = c(0,
                 # tmp[tmp$team == "baseline_mean" &
                 #       tmp$scenario == "across-subject", ]$m.err + .02, 
                 tmp[tmp$team == "baseline_random" &
                       tmp$scenario == "across-subject", ]$m.err - .02),
           x = .9,
           size = 1,
           colour = "grey40") +
  
  annotate(geom = "text",
           y = 2,
           x = .75,
           label = "above chance",
           angle = 90,
           vjust = -1,
           colour = "grey40") +
  
  # add teams
  geom_line(data = filter(tmp, 
                          !str_detect(team, "baseline")), 
               alpha = .5, 
            size = 2) +
  
  ggrepel::geom_text_repel(data = filter(tmp, 
                                         !str_detect(team, "baseline"), 
                                         scenario == "across-time"), 
                           aes(label = team, 
                               col = team),
                           hjust = 1,
                           nudge_x = 1,
                           force = 2) + 
  scale_x_discrete(expand = expansion(mult = c(.2, .2))) +
  #scale_y_continuous(expand = expansion(mult = c(0, 0))) +
  ylab('mean absolute prediction error') +
  scale_y_reverse(expand = expansion(mult = c(0, 0))) +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  
  ggrepel::geom_text_repel(data = filter(tmp, 
                                         team == "team 5",
                                         scenario == "across-subject"),
                           aes(label = "italic('more')~accurate~across~subjects~than~time",
                               col = team),
                           nudge_y = .3,
                           nudge_x = .9,
                           parse = T) +
  
  ggrepel::geom_text_repel(data = filter(tmp, 
                                         team == "team 8",
                                         scenario == "across-subject"),
                           aes(label = "italic('less')~accurate~across~subjects~than~time",
                               col = team),
                           nudge_y = -.3,
                           nudge_x = .9,
                           parse = T)
```

# Examine impact of simulated physiological randomness 
Create SimPhysioAnalyses function
```{r}
SimPhysioAnalyses <- function(t){
# filter data
d <- DF %>% 
filter(grepl(t, team)) %>%
# group_by(team, scenario) %>% 
# sample_n(size = 500) %>% 
# ungroup() %>% 
mutate(team = factor(team))

# distinguish between real vs. simulated physiology
# d <- d %>% 
#   mutate(phys.real = 
#            if_else(grepl('random', team),
#                    true = 'simulated',
#                    false = 'real'))

# fit model
m <- lmer(error ~ scenario * team +
(1 | sub) +
(1 | vid),
data = d)

# run ANOVA and perform contrasts
# m.aov <- m %>% 
#   anova() %>% 
#   mutate(team = t)

m.emm <- m %>% 
emmeans(pairwise ~ team | scenario) %>% 
.$contrasts %>% 
as.data.frame %>% 
filter(grepl(t, contrast)) %>% 
mutate(team = t, 
contrast = str_sub(contrast,
start = 1, end = 6))

# compile results into list
return(m.emm)
}
```

Run function
```{r}
t.list <- c("team 3", "team 4", "team 9")

sim.results <- sapply(FUN = SimPhysioAnalyses,
X = t.list,
simplify = F)

rm(t, t.list, d, m, m.emm)

saveRDS(sim.results, 'sim.results.Rds')
# sim.results <- readRDS('sim.results.Rds')
```

# Figure 2
Create figure dataframe
```{r}
# create dataframe containing contrasts
DF.f2.contrasts <- bind_rows(sim.results) %>% 
select(scenario, estimate, z.ratio, p.value, team) %>% 
mutate(p.value = stars.pval(p.value),
estimate = round(estimate, 2))

# create dataframe containing baseline performance
DF.f2.baselines <- DF %>% 
filter(team == "baseline_mean" |
team == "baseline_random") %>% 
group_by(team, scenario) %>% 
summarise(m.err = mean(error)) %>% 
pivot_wider(names_from = team, 
values_from = m.err)

# create dataframe containing team performance
DF.f2.teams <- DF %>% 
filter(grepl("team 3", team) |
grepl("team 4", team) |
grepl("team 9", team)) %>% 
group_by(team, scenario) %>% 
summarise(m.err = mean(error)) %>% 
mutate(rand = 
if_else(grepl("random", team),
true = "random",
false = "real"),
team = str_sub(team,
start = 1, end = 6))

# merge dataframes
DF.f2 <- 
full_join(x = DF.f2.teams,
y = DF.f2.baselines) %>% 
full_join(x = .,
y = DF.f2.contrasts) %>% 
mutate(scenario = 
factor(scenario,
labels = 
c('across-time validation',
'across-subject validation',
'across-emotion validation' ,
'across-induction validation')
),
scenario =
factor(scenario,
levels = c('across-subject validation',
'across-emotion validation' ,
'across-induction validation',
'across-time validation')),
team =
factor(team,
levels = c("team 9", "team 4", "team 3"))
)

rm(DF.f2.baselines, DF.f2.contrasts, DF.f2.teams)
```

Descriptives
```{r}
DF.f2 %>% 
filter(estimate != 0.00) %>% 
group_by(scenario) %>% 
summarise(min.diff = min(estimate),
max.diff = max(estimate),
min.z = min(z.ratio),
max.z = max(z.ratio)) %>% 
View()
```


```{r}
ggplot(DF.f2,
aes(x = m.err,
y = team,
fill = rand)) +
facet_wrap(~ scenario, 
nrow = 4, 
ncol = 1) +

# add team performance bars
geom_bar(stat= "identity",
position = 'dodge') +

# add lines for baseline_mean performance
# geom_vline(linetype = 'dashed',
#            aes(xintercept = baseline_mean),
#            colour = '#3366FF') +

# add summary statistic text
geom_text(aes(x = 1.75,
label = paste0(estimate %>% 
format(nsmall = 2), 
p.value)
),
size = 3) +
ylab('model developers') +
xlab('mean absolute prediction error') +
scale_x_continuous(expand = 
expansion(mult = c(0, .05))
) +
scale_fill_manual(values=c("grey",
"#C0C0F0")) +
theme(legend.position = "top",
legend.title = element_blank())

```



# JUNK
Old code for formatting contrasts
```{r}
mutate(contrast = str_sub(contrast,
start = 9, end = 13),
contrast.type = paste0(scenario, "_", contrast),
p.value = stars.pval(p.value),
summary = 
paste0("M_dif = ",
round(estimate, 2),
", z = ",
round(z.ratio, 2),
p.value)
) %>% 
select(contrast.type, summary) %>% 
pivot_wider(names_from = contrast.type,
values_from = summary)
```


```{r}
DF.f1a <- DF.f1 %>% 
filter(team == "baseline_mean" |
team == "baseline_random" |
team == "baseline_tailored")

DF.f1b <- DF.f1 %>% 
filter(team != "baseline_mean",
team != "baseline_random",
team != "baseline_tailored")

mutate(`Pr(>F)` = stars.pval(`Pr(>F)`),
summary = paste0("F(",
NumDF, ",",
round(DenDF, 2), 
") = ",
round(`F value`, 2),
`Pr(>F)`)
)
```


Run function
```{r}
t.list <- c('team 3', 'team 4')

results <- sapply(FUN = MainAnalyses,
X = t.list)

t <- 'team 4'
m.aov <- results[[2]][["m.aov"]] %>% 
mutate(team = t)


m.emm <- results[[2]][["m.emm"]]$contrasts %>%  
as.data.frame %>%
filter(grepl('team 4', contrast)) %>% 
mutate(team = t, 
contrast = str_sub(contrast,
start = 9, end = 13))



tmp <- as.dataframe(team = t,

# ANOVA output
aov.s = m.aov['scenario', 'summary'],
aov.t = m.aov['team', 'summary'],
aov.st = m.aov['scenario:team', 'summary']

# contrast output
)
```


Plot figure alternative
```{r}
ggplot(DF.f1,
aes(x = team,
y = m.err,
fill = scenario)) +
geom_bar(stat= 'identity',
position = 'dodge')
```

Plot figure alternative 2
```{r}
tmp <- DF %>% 
filter(team != "team 3 (random)",
team != "team 4 (random)",
team != "team 9 (random)",
team != "baseline_mean",
team != "baseline_random",
team != "baseline_tailored")

tmp2 <- DF %>% 
filter(team == "baseline_mean" |
team == "baseline_random" |
team == "baseline_tailored") %>% 
group_by(scenario, team) %>% 
summarise(error = mean(error))

ggplot(data = tmp,
aes(y = team,
x = error,
colour = scenario)) +
stat_summary(fun.data = mean_se,
geom = "errorbar") + 
geom_vline(data = tmp2 %>% 
filter(team == 'baseline_mean'),
linetype = 'dashed',
aes(xintercept = error,
colour = scenario))




ggplot(tmp,
aes(x = team,
y = error,
colour = scenario)) +
stat_summary(fun.y = mean,
fun.ymin = function(x) mean(x) - sd(x), 
fun.ymax = function(x) mean(x) + sd(x),
geom = "errorbar")
```
