---
title: "EPiC_23a Data Processing Script"
author: "Nicholas Coles"
date: "`r Sys.Date()`"
output: word_document
editor_options: 
  chunk_output_type: console
---

# Set environment
```{r}
# clear environment
rm(list = ls())

# install (if necessary) and load packages
# function written by stevenworthington 
Ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, 'Package'])]
  
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  
  sapply(pkg, require, character.only = TRUE)
}

# vector of necessary packages
packages <- c('tidyverse', 'lme4',
              'lmerTest', 'emmeans',
              'gtools', 'gridExtra')

# using vector of packages, call ipak function
Ipak(packages)

# delete vestigial
rm(packages, Ipak)

# turn scientific notation off
options(scipen = 999)

# set seed to year of Nicholas' favorite [unfinished] album, SMiLE
set.seed(1967)

# allow for Type III Sums of Square
options(contrasts = c('contr.sum', 'contr.poly'))

# set plotting theme
theme_set(theme_classic())

# free up computational resources
gc()
memory.limit(size = 25000)
```

# Load data
```{r}
DF <- readRDS('C:/Users/Nicholas Coles work/Desktop/local/DF.rds') 
```

# Create the mega-model
```{r eval = F}
tmp <- DF %>% 
  filter(team != "team 4 (random)",
         team != "team 9 (random)",
         team != "team 3 (random)",
         team != "baseline_tailored") %>% 
  
  # extract info
  mutate(source = 
           if_else(condition = grepl("team", team),
                   true = "team",
                   false = "baseline"),
         team.name = 
           if_else(source == "team",
                   true = team,
                   false = NA),
         baseline.type = 
           if_else(team == "baseline_random",
                   true = "random",
                   false = 
                     if_else(team == "baseline_mean",
                             true = "mean",
                             false = NA)
           )
  ) %>% 
  
  # refactor
  mutate(source = factor(source),
         team.name = factor(team.name),
         baseline.type = factor(baseline.type))

mega.m <- 
  lmer(error ~ source : baseline.type : team.name : scenario +
         (1 | sub) + (1 | vid),
       data = tmp)
```


# Examine error in  teams' submission
Create MainAnalyses function
```{r}
MainAnalyses <- function(t){
  # filter data
  d <- DF %>% 
    filter(team == "baseline_random" |
             team == "baseline_mean" |
             team == t) %>%
    #group_by(scenario) %>% 
    #sample_n(size = 1000) %>% 
    #ungroup() %>% 
    mutate(team = factor(team))
  
  # fit model
  m <- lmer(error ~ scenario * team +
              (1 | sub_orig) +
              (1 | vid_orig),
            data = d)
  
  # run ANOVA and perform contrasts
  # m.aov <- m %>% 
  #   anova() %>% 
  #   mutate(team = t)
  
  m.emm <- m %>% 
    emmeans(pairwise ~ team | scenario) %>% 
    .$contrasts %>% 
    as.data.frame %>% 
    filter(grepl(t, contrast)) %>% 
    mutate(team = t, 
           contrast = str_sub(contrast,
                              start = 9, end = 13))
  
  # compile results into list
  return(m.emm)
}
```

Run function
```{r}
t.list <- c("team 1", "team 2", "team 3",
            "team 4", "team 5", "team 6",
            "team 7", "team 8", "team 9",
            "team 10", "team 11", "team 12")

results <- sapply(FUN = MainAnalyses,
                  X = t.list,
                  simplify = F)

rm(t, t.list, d, m, m.emm)

saveRDS(results, 'results.Rds')

#results <- readRDS('C:/Users/Nicholas Coles work/Desktop/local/results.Rds')
```

# Figure 1

## Prep dataframe
```{r}
# open results
#results <- readRDS('results.Rds')

# create dataframe containing contrasts
DF.f1.contrasts <- bind_rows(results) %>%
  select(contrast:estimate, z.ratio, p.value, team) %>%
  mutate(p.value = stars.pval(p.value),
         p.value = if_else(
           condition = p.value == " ",
           true = "n.s.",
           false = if_else(
             condition = p.value == "*",
             true = "*",
             false = " "
           )
         ),
         estimate = round(estimate, 2)) %>%
  pivot_wider(names_from = contrast,
              values_from = c(estimate, z.ratio, p.value))

# create dataframe containing baseline performance
DF.f1.baselines <- DF %>%
  filter(team == "baseline_mean" |
           team == "baseline_random") %>%
  group_by(team, scenario) %>%
  summarise(m.err = mean(error)) %>%
  pivot_wider(names_from = team,
              values_from = m.err)

# create dataframe containing team performance
DF.f1.teams <- DF %>%
  filter(team != "team 3 (random)",
         team != "team 4 (random)",
         team != "team 9 (random)",
         team != "baseline_mean",
         team != "baseline_random",
         team != "baseline_tailored") %>%
  group_by(team, scenario) %>%
  summarise(m.err = mean(error))

# merge dataframes
DF.f1 <-
  full_join(x = DF.f1.contrasts,
            y = DF.f1.baselines) %>%
  full_join(x = .,
            y = DF.f1.teams) %>%
  mutate(
    scenario = factor(scenario,
                      labels = c('time',
                                 'subject',
                                 'emotion' ,
                                 'induction')),
    scenario = factor(scenario,
                      levels = c('subject',
                                 'emotion' ,
                                 'induction',
                                 'time')),
    team = factor(team,
                  levels = paste0('team ',
                                  seq(12, 1)))
  )

rm(DF.f1.baselines, DF.f1.contrasts, DF.f1.teams)
```

## Plot figure
```{r}
fig1 <-
  ggplot(DF.f1,
         aes(x = m.err,
             y = team)) +
  
  # facet plot
  facet_wrap(~ scenario,
             nrow = 4,
             ncol = 1) +
  # reverse x scale
  scale_x_reverse(expand =
                    expansion(mult = c(0, .05))) +
  
  # create lollipop plot
  geom_segment(aes(y = team, 
                   yend = team, 
                   x = 2.75, 
                   xend = m.err),
               colour = "grey90",
               linewidth = 1.5) +
  geom_point(aes(color = team),
             size = 4) +
  
  # add lines for baseline_random and baseline_mean performance
  geom_vline(linetype = 'dotted',
             aes(xintercept = baseline_random),
             colour = '#3366FF') +
  geom_vline(linetype = 'dashed',
             aes(xintercept = baseline_mean),
             colour = '#3366FF') +
  
  # add summary statistic text (to show where models significantly differed from mean baseline)
  geom_text(aes(x = baseline_mean - .05,
                label = paste0(p.value__mean)),
            size = 3,
            hjust = 0) +
  ylab('model developers') +
  xlab('mean absolute prediction error') +
  theme(legend.position = "none")

#rm(DF.f1)
```

# Figure 2
```{r}
fig2.list <- 
  lapply(X = unique(DF.f1$team),
         FUN = function(x){
           # identify if case is the team you want to highlight
           d <- DF.f1 %>%
             mutate(highlight =
                      if_else(team == x,
                              1,
                              0),
                    highlight = factor(highlight))
           
           # create plot
           g <- ggplot(d,
                       aes(y = m.err,
                           x = scenario,
                           group = team,
                           colour = team,
                           size = highlight)) +
             
             # plot non-highlighted lines in grey
             geom_line(data = d %>%
                         filter(highlight == 0),
                       colour = "grey90",
                       linewidth = 1) +
             
             # plot all lines, but then remove the non-highlighted ones
             geom_line() +
             scale_size_manual(values = c(NA, 3)) +
             
             # clean up aesthetics
             theme(legend.position = "none",
                   axis.title = element_blank(),
                   ) +
             scale_y_reverse() +
             
             # conditional formatting to manually make it look like a facet
             {if(x != "team 10" & 
                 x != "team 11" &
                 x != "team 12")
               theme(axis.title.x = element_blank(),
                     axis.text.x = element_blank())} +
             
             {if(x != "team 1" & 
                 x != "team 4" &
                 x != "team 7" &
                 x != "team 10")
               theme(axis.title.y = element_blank(),
                     axis.text.y = element_blank())}

           # return plot
           return(g)
         }
       )


fig2 <- grid.arrange(grobs = fig2.list,
                     left = "mean absolute prediction error",
                     bottom = "validation")
```

For Stan
```{r}
library('cowplot')
plot_grid(fig1, NULL, fig2,
          rel_widths = c(1, .1, 1),
          labels = c('A', '', 'B'),
          ncol = 3)
```

# Figure 3
This figure could be cool because it could demonstrate that the test data has the same differentiation as the Sharma dataset. Test change

Could be more trouble than it's worth though.

You could imagine this being the A part of a spirit figure.

Particpiants encountered 4 inductions of two affective states while providing moment-to-moment measures of valence and arousal

Data are split for four types of validation

```{r}
# add video label
DF <- DF %>% 
  mutate(`targeted emotion` = 
           if_else(
             vid_orig == 1 | 
               vid_orig == 2,
             true = "amusing",
             false = 
               if_else(
                 vid_orig == 3 |
                   vid_orig == 4,
                 true = "boring",
                 false = 
                   if_else(
                     vid_orig == 5 |
                       vid_orig == 6,
                     true = "relaxing",
                     false = "scary")
               )
             )
         )

# create participant level summaries
f0.dfa <- DF %>% 
  filter(team == "baseline_mean",
         scenario == "scenario_4") %>% 
  select(sub_orig, vid_orig, `targeted emotion`, time, scenario, fold, outcome, test) %>% 
  pivot_wider(names_from = 'outcome',
              values_from = 'test') %>% 
  group_by(sub_orig, `targeted emotion`) %>% 
  summarise(m.val = mean(valence),
            m.aro = mean(arousal))

# create video level summaries
f0.dfb <- DF %>% 
  filter(team == "baseline_mean",
         scenario == "scenario_4") %>% 
  select(sub_orig, vid_orig, `targeted emotion`, time, scenario, fold, outcome, test) %>% 
  pivot_wider(names_from = 'outcome',
              values_from = 'test') %>% 
  group_by(`targeted emotion`) %>% 
  summarise(m.val = mean(valence),
            sd.val = sd(valence),
            m.aro = mean(arousal),
            sd.aro = sd(arousal))

ggplot(data = f0.dfa,
       aes(x = m.val,
           y = m.aro,
           fill = `targeted emotion`
           )
       ) +
  stat_ellipse(geom = "polygon",
               alpha = .2) +
  geom_point(aes(color = `targeted emotion`)) +
  ylab('mean valence reports') +
  xlab('mean arousal reports')
```

# Figure 0b




# New idea for Figure 1
```{r}
tmp <- DF %>% 
  group_by(team, scenario) %>% 
  summarise(m.err = mean(error)) %>% 
  filter(team != "team 9 (random)",
         team != "team 4 (random)",
         team != "team 3 (random)") %>% 
  mutate(scenario = 
           factor(scenario,
                  labels = 
                    c('across-time',
                      'across-subject',
                      'across-emotion' ,
                      'across-induction')),
         scenario =
           factor(scenario,
                  levels = c('across-subject',
                             'across-emotion' ,
                             'across-induction',
                             'across-time'))
  )

# ggplot(tmp,
#        aes(x = scenario,
#            y = m.err,
#            group = team,
#            colour = team)) +
#   stat_summary(geom = "line",
#                size = 1) +
#   scale_y_reverse() +
#   scale_x_discrete(position = "top") 
```

Next steps for this:
* Annotate Team 5 and 8
* Higher accuracy than mean baseline model (doesn't look great)
* Higher accuracy than random baseline model (doesn't look great)
* Re-consider ordering of teams

Thought about figure: we could hypothetically show people the distribution of the data set


Thought:
Red team.

The models appear to be capturing something. Whether it's a true link with physiology, we cannot say for sure.


Core affect may not merely be physiological

Are links between physiology and core affect 

##  Mikes's code                       
```{r}
g <- ggplot(tmp,
       aes(x = scenario,
           y = m.err,
           group = team,
           colour = team)) +
  
  # create baseline areas
  geom_area(data = filter(tmp, 
                          str_detect(team, "baseline_random")), 
            colour = NA,
            fill = "gray85") +
  geom_area(data = filter(tmp, 
                          str_detect(team, "baseline_mean")), 
            colour = NA,
            fill = "grey95") +
  
  # label  baseline areas
  ## baseline_mean
  annotate(geom = "line",
           y = c(0 + .02, 
                 tmp[tmp$team == "baseline_mean" &
                       tmp$scenario == "across-subject", ]$m.err - .02),
           x = .75,
           size = 1,
           colour = "grey60"
           ) +
  
  annotate(geom = "text",
           y = tmp[tmp$team == "baseline_mean" &
                     tmp$scenario == "across-subject", ]$m.err / 2,
           x = .75,
           label = "above relying on sample averages",
           colour = "grey60",
           angle = 90,
           vjust = -1)

g <- g +
  ## baseline_random
  annotate(geom = "line",
           y = c(0,
                 # tmp[tmp$team == "baseline_mean" &
                 #       tmp$scenario == "across-subject", ]$m.err + .02, 
                 tmp[tmp$team == "baseline_random" &
                       tmp$scenario == "across-subject", ]$m.err - .02),
           x = .9,
           size = 1,
           colour = "grey40") +
  
  annotate(geom = "text",
           y = 2,
           x = .75,
           label = "above chance",
           angle = 90,
           vjust = -1,
           colour = "grey40") +
  
  # add teams
  geom_line(data = filter(tmp, 
                          !str_detect(team, "baseline")), 
               alpha = .5, 
            size = 2) +
  
  ggrepel::geom_text_repel(data = filter(tmp, 
                                         !str_detect(team, "baseline"), 
                                         scenario == "across-time"), 
                           aes(label = team, 
                               col = team),
                           hjust = 1,
                           nudge_x = 1,
                           force = 2) + 
  scale_x_discrete(expand = expansion(mult = c(.2, .2))) +
  #scale_y_continuous(expand = expansion(mult = c(0, 0))) +
  ylab('mean absolute prediction error') +
  scale_y_reverse(expand = expansion(mult = c(0, 0))) +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  
  ggrepel::geom_text_repel(data = filter(tmp, 
                                         team == "team 5",
                                         scenario == "across-subject"),
                           aes(label = "italic('more')~accurate~across~subjects~than~time",
                               col = team),
                           nudge_y = .3,
                           nudge_x = .9,
                           parse = T) +
  
  ggrepel::geom_text_repel(data = filter(tmp, 
                                         team == "team 8",
                                         scenario == "across-subject"),
                           aes(label = "italic('less')~accurate~across~subjects~than~time",
                               col = team),
                           nudge_y = -.3,
                           nudge_x = .9,
                           parse = T)
```

# Examine impact of simulated physiological randomness 
Create SimPhysioAnalyses function
```{r}
SimPhysioAnalyses <- function(t){
# filter data
d <- DF %>% 
filter(grepl(t, team)) %>%
# group_by(team, scenario) %>% 
# sample_n(size = 500) %>% 
# ungroup() %>% 
mutate(team = factor(team))

# distinguish between real vs. simulated physiology
# d <- d %>% 
#   mutate(phys.real = 
#            if_else(grepl('random', team),
#                    true = 'simulated',
#                    false = 'real'))

# fit model
m <- lmer(error ~ scenario * team +
(1 | sub) +
(1 | vid),
data = d)

# run ANOVA and perform contrasts
# m.aov <- m %>% 
#   anova() %>% 
#   mutate(team = t)

m.emm <- m %>% 
emmeans(pairwise ~ team | scenario) %>% 
.$contrasts %>% 
as.data.frame %>% 
filter(grepl(t, contrast)) %>% 
mutate(team = t, 
contrast = str_sub(contrast,
start = 1, end = 6))

# compile results into list
return(m.emm)
}
```

Run function
```{r}
t.list <- c("team 3", "team 4", "team 9")

sim.results <- sapply(FUN = SimPhysioAnalyses,
X = t.list,
simplify = F)

rm(t, t.list, d, m, m.emm)

saveRDS(sim.results, 'sim.results.Rds')
# sim.results <- readRDS('sim.results.Rds')
```

# Figure 2
Create figure dataframe
```{r}
# create dataframe containing contrasts
DF.f2.contrasts <- bind_rows(sim.results) %>% 
select(scenario, estimate, z.ratio, p.value, team) %>% 
mutate(p.value = stars.pval(p.value),
estimate = round(estimate, 2))

# create dataframe containing baseline performance
DF.f2.baselines <- DF %>% 
filter(team == "baseline_mean" |
team == "baseline_random") %>% 
group_by(team, scenario) %>% 
summarise(m.err = mean(error)) %>% 
pivot_wider(names_from = team, 
values_from = m.err)

# create dataframe containing team performance
DF.f2.teams <- DF %>% 
filter(grepl("team 3", team) |
grepl("team 4", team) |
grepl("team 9", team)) %>% 
group_by(team, scenario) %>% 
summarise(m.err = mean(error)) %>% 
mutate(rand = 
if_else(grepl("random", team),
true = "random",
false = "real"),
team = str_sub(team,
start = 1, end = 6))

# merge dataframes
DF.f2 <- 
full_join(x = DF.f2.teams,
y = DF.f2.baselines) %>% 
full_join(x = .,
y = DF.f2.contrasts) %>% 
mutate(scenario = 
factor(scenario,
labels = 
c('across-time validation',
'across-subject validation',
'across-emotion validation' ,
'across-induction validation')
),
scenario =
factor(scenario,
levels = c('across-subject validation',
'across-emotion validation' ,
'across-induction validation',
'across-time validation')),
team =
factor(team,
levels = c("team 9", "team 4", "team 3"))
)

rm(DF.f2.baselines, DF.f2.contrasts, DF.f2.teams)
```

Descriptives
```{r}
DF.f2 %>% 
filter(estimate != 0.00) %>% 
group_by(scenario) %>% 
summarise(min.diff = min(estimate),
max.diff = max(estimate),
min.z = min(z.ratio),
max.z = max(z.ratio)) %>% 
View()
```


```{r}
ggplot(DF.f2,
aes(x = m.err,
y = team,
fill = rand)) +
facet_wrap(~ scenario, 
nrow = 4, 
ncol = 1) +

# add team performance bars
geom_bar(stat= "identity",
position = 'dodge') +

# add lines for baseline_mean performance
# geom_vline(linetype = 'dashed',
#            aes(xintercept = baseline_mean),
#            colour = '#3366FF') +

# add summary statistic text
geom_text(aes(x = 1.75,
label = paste0(estimate %>% 
format(nsmall = 2), 
p.value)
),
size = 3) +
ylab('model developers') +
xlab('mean absolute prediction error') +
scale_x_continuous(expand = 
expansion(mult = c(0, .05))
) +
scale_fill_manual(values=c("grey",
"#C0C0F0")) +
theme(legend.position = "top",
legend.title = element_blank())

```



# JUNK
Old code for formatting contrasts
```{r}
mutate(contrast = str_sub(contrast,
start = 9, end = 13),
contrast.type = paste0(scenario, "_", contrast),
p.value = stars.pval(p.value),
summary = 
paste0("M_dif = ",
round(estimate, 2),
", z = ",
round(z.ratio, 2),
p.value)
) %>% 
select(contrast.type, summary) %>% 
pivot_wider(names_from = contrast.type,
values_from = summary)
```


```{r}
DF.f1a <- DF.f1 %>% 
filter(team == "baseline_mean" |
team == "baseline_random" |
team == "baseline_tailored")

DF.f1b <- DF.f1 %>% 
filter(team != "baseline_mean",
team != "baseline_random",
team != "baseline_tailored")

mutate(`Pr(>F)` = stars.pval(`Pr(>F)`),
summary = paste0("F(",
NumDF, ",",
round(DenDF, 2), 
") = ",
round(`F value`, 2),
`Pr(>F)`)
)
```


Run function
```{r}
t.list <- c('team 3', 'team 4')

results <- sapply(FUN = MainAnalyses,
X = t.list)

t <- 'team 4'
m.aov <- results[[2]][["m.aov"]] %>% 
mutate(team = t)


m.emm <- results[[2]][["m.emm"]]$contrasts %>%  
as.data.frame %>%
filter(grepl('team 4', contrast)) %>% 
mutate(team = t, 
contrast = str_sub(contrast,
start = 9, end = 13))



tmp <- as.dataframe(team = t,

# ANOVA output
aov.s = m.aov['scenario', 'summary'],
aov.t = m.aov['team', 'summary'],
aov.st = m.aov['scenario:team', 'summary']

# contrast output
)
```


Plot figure alternative
```{r}
ggplot(DF.f1,
aes(x = team,
y = m.err,
fill = scenario)) +
geom_bar(stat= 'identity',
position = 'dodge')
```

Plot figure alternative 2
```{r}
tmp <- DF %>% 
filter(team != "team 3 (random)",
team != "team 4 (random)",
team != "team 9 (random)",
team != "baseline_mean",
team != "baseline_random",
team != "baseline_tailored")

tmp2 <- DF %>% 
filter(team == "baseline_mean" |
team == "baseline_random" |
team == "baseline_tailored") %>% 
group_by(scenario, team) %>% 
summarise(error = mean(error))

ggplot(data = tmp,
aes(y = team,
x = error,
colour = scenario)) +
stat_summary(fun.data = mean_se,
geom = "errorbar") + 
geom_vline(data = tmp2 %>% 
filter(team == 'baseline_mean'),
linetype = 'dashed',
aes(xintercept = error,
colour = scenario))




ggplot(tmp,
aes(x = team,
y = error,
colour = scenario)) +
stat_summary(fun.y = mean,
fun.ymin = function(x) mean(x) - sd(x), 
fun.ymax = function(x) mean(x) + sd(x),
geom = "errorbar")
```
